Analyzing Inflation Dynamics in Canada During and After the COVID-19 Pandemic: A Deep Dive into Consumer Price Index Movements from 2019 to 2023

### 2024-04-21 outputs/paper.qmd
Update

## Data Analysis
This paper utilizes data from Statistics Canada's Consumer Price Index (CPI), focusing on a subset of six variables that represent diverse categories of consumer spending: food, household essentials, clothing, gasoline, healthcare, and recreation. These categories were carefully selected to encompass a broad spectrum of goods and services, providing a comprehensive sight of consumer price changes during significant economic periods. The analysis covers the period from 2019 to 2023, a timeline strategically chosen to encompass the main phases of the COVID-19 pandemic as well as the subsequent post-pandemic economic environment.

@@

```
@tbl-1 offers a numerical perspective on the trends highlighted in Figure 1, providing the exact annual values that reflect the differential inflation rates for Food and All-items. It is clear from the table that the Food Value consistently increased at a higher rate than the All-items Value each consecutive year. Notably, between 2022 and 2023, the Food Value rose from 171.5 to 184.3, while the All-items Value exhibited a more moderate climb from 151.2 to 157.1. This quantifiable data underscores the disproportionate inflationary pressures on food items.

>>>>

## Dataset

The Dataset were derived from Statistics Canada's Consumer Price Index (CPI), capturing key statistics on Canadian Consumaer Price Index Value of different items encompassing from 2019 to 2023.

@tbl-7 shows the first 5 rows in the dataset, REF_DATE stands for the time that the data was gathered, products_and_product_groups stands for items category, value stands for CPI value.

```{r}
#| label: tbl-7
#| tbl-cap: First 5 Rows in Mental Health Indicators Dataset 
#| echo: false

kable(head(product_data,5))
```

## Data Analysis
To analyze the data , I worked on R [@citeR], using the tidyverse [@tidyverse], [@arrow], and gitcreds [@gitcreds] packages. Tables in the report were created with kableExtra [@kableExtra] and figures were created with ggplot2 which was built in tidyverse [@tidyverse].

This paper utilizes data from Statistics Canada's Consumer Price Index (CPI), focusing on a subset of six variables that represent diverse categories of consumer spending: food, household essentials, clothing, gasoline, healthcare, and recreation. These categories were carefully selected to encompass a broad spectrum of goods and services, providing a comprehensive sight of consumer price changes during significant economic periods. The analysis covers the period from 2019 to 2023, a timeline strategically chosen to encompass the main phases of the COVID-19 pandemic as well as the subsequent post-pandemic economic environment.

@@

```
@tbl-1 offers a numerical perspective on the trends highlighted in @fig-1, providing the exact annual values that reflect the differential inflation rates for Food and All-items. It is clear from the table that the Food Value consistently increased at a higher rate than the All-items Value each consecutive year. Notably, between 2022 and 2023, the Food Value rose from 171.5 to 184.3, while the All-items Value exhibited a more moderate climb from 151.2 to 157.1. This quantifiable data underscores the disproportionate inflationary pressures on food items.
